---
title: "Sales Forecast"
author: "Yu Yuen Hern"
date: "`r Sys.Date()`"
output: html_notebook
runtime: shiny
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r, echo=F}
library(readxl)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(stats)

DATASET_PATH <- "dataset/Data.xlsx"
```

## Read dataset
```{r}
df <- readxl::read_xlsx(DATASET_PATH, sheet = "Data")
df
```

## Check for missing values
```{r}
sum(is.na(df$`Sales Amt`))
```

```{r}
df$`Sales Amt`[is.na(df$`Sales Amt`)] = 0
sum(is.na(df$`Sales Amt`))
```

## Plot sales over time
```{r}
p <- ggplot(df, aes(x=`PO Date`, y=`Sales Amt`)) + geom_line()
p
```

## Holt Winters

As Holt-Winters function require time series `ts` object and `ts` object only accepts evenly spaced data, we aggregate the daily data into monthly data.
```{r}
df <- df[c("Year", "Month", "Sales Amt")]
df
```

Aggregate the data
```{r}
df <- df %>% group_by(Year, Month) %>% summarise(Monthly_Sales=sum(`Sales Amt`))
df
```

Split the dataset into train (2019 and 2020) and test (2021)
```{r}
df_train <- subset(df, Year <= 2020)
df_test <- subset(df, Year == 2021)
```

Overall plot of data
```{r}
dfts <- df$Monthly_Sales %>% ts(start = c(2019, 1), end = c(2021, 12), frequency = 12)
plot(dfts)
```

Convert the data into `ts` object
```{r}
dfts_train <- df_train$Monthly_Sales %>% ts(start = c(2019, 1), end = c(2020, 12), frequency = 12)
dfts_train
```

Perform Holt-Winters fitting
```{r}
hw1 <- HoltWinters(dfts_train)
hw1
```

Perform Holt-Winters prediction and overlay forecast on original data
```{r}
hw1.pred <- predict(hw1, n.ahead = 12, prediction.interval = TRUE, level = 0.95)
plot(hw1, predicted.values = hw1.pred, ylab = "Sales Amt")
```

Perform Holt-Winters fitting with multiplicative seasonality
```{r}
hw2 <- HoltWinters(dfts_train, seasonal = "multiplicative")
hw2.pred <- predict(hw2, n.ahead = 12, prediction.interval = TRUE, level = 0.95)
plot(hw2, predicted.values = hw2.pred, ylab = "Sales Amt")
```

## Using forecast library
```{r}
library(forecast)
```

Plot hw1 model with forecast library
```{r}
hw1_for <- forecast(hw1, h = 12, level = 0.95)
plot(hw1_for)
lines(hw1_for$fitted, lty=2, col="red")
```

Plot hw2 model with forecast library
```{r}
hw2_for <- forecast(hw2, h = 12, level = 0.95)
plot(hw2_for)
lines(hw2_for$fitted, lty=2, col="red")
```

## Evaluate

Firstly we want to check if there are any autocorrelation in residue.
```{r}
hw1_acf <- acf(hw1_for$residuals, lag.max = 12, na.action=na.pass)
```

There is no autocorrelation as the values are all below blue line.

Secondly, we want to find out if the variable are independent using Ljung-Box test.
```{r}
Box.test(hw1_for$residuals, type="Ljung-Box")
```

The p-value from Ljung Box Test is 0.1428 for lag=1, hence we fail reject the hypothesis. The variables are independent.

Thirdly, we check if the residuals are normally distributed.
```{r}
hist(hw1_for$residuals)
```

The histogram is normally distributed.

And lastly, we check the MAPE of both Holt-Winters model.
```{r}
mape <- function(actual,pred){
  mape <- mean(abs((actual - pred)/actual))*100
  return (mape)
}
```

```{r}
mape_hw1 <- mape(df_test$Monthly_Sales, hw1.pred[,1])
mape_hw2 <- mape(df_test$Monthly_Sales, hw2.pred[,1])
cat(mape_hw1, mape_hw2)
```
MAPE for HW1 is 18.76 while HW2 is 17.79

## Forecast 36 months (12 months of dfts_test + 24 months as requested)
```{r}
dfts_test <- df_test$Monthly_Sales %>% ts(start = c(2021, 1), end = c(2021, 12), frequency = 12)

hw1_for <- forecast(hw1, h = 36, level = 0.95)
plot(hw1_for, ylim = c(0,2.3e+7))
lines(dfts, lty = 2, col = "red")
legend(x = "topright", legend=c("Forecast", "Fitted", "Original"), col=c("blue", "black", "red"), lty=1:2, cex = 0.6)
```

Original series is in red, fitted values in black and forecast values in blue.
